Thermodynamic response functions in a cell fluid model

### 2024-09-01 main_morse_response_functions.tex
write two equivalent expressions for alpha_P suitable for two forms of eos

	\subsection{Thermal expansion coefficient}
	The thermal expansion coefficient is defined by
	\begin{equation}
		\alpha_P = \frac{1}{V}\left(\frac{\partial V}{\partial T}\right)_{P,\langle N\rangle}
	\end{equation}
	We rewrite $\alpha_P$ into a form suitable for the equation of state~\eqref{eq:eosMT}
	\begin{eqnarray*}
		\alpha_P & = & \frac{1}{V}\left(\frac{\partial V}{\partial T}\right)_{P,\langle N\rangle}
		\\
		& = & -\frac{1}{\rho} \left(\frac{\partial \rho}{\partial T}\right)_{P, N}
		\\
		& = & -\frac{1}{\rho} \left(\frac{\partial \rho}{\partial T}\right)_{\mu} 
		 - \frac{1}{\rho} \left(\frac{\partial \rho}{\partial \mu}\right)_{T}
		 \left(\frac{\partial \mu}{\partial T}\right)_{P, N}.
	\end{eqnarray*}
	Applying the cyclic relation
	\begin{equation*}
		\left(\frac{\partial \mu}{\partial T}\right)_{P, N}
		\left(\frac{\partial T}{\partial P}\right)_{\mu, N}
		\left(\frac{\partial P}{\partial \mu}\right)_{T, N}
		= -1
	\end{equation*}
	we get
	\begin{eqnarray*}
		\left(\frac{\partial \mu}{\partial T}\right)_{P, N} & = & 
		- \left(\frac{\partial P}{\partial T}\right)_{\mu}
		\left(\frac{\partial P}{\partial \mu}\right)^{-1}_{T}.
	\end{eqnarray*}
	Substituting the result into the expression for $\alpha_P$, we arrive at the final formula for the thermal expansion coefficient
	\begin{equation}
		\alpha_P = -\frac{1}{\rho}\left(\frac{\partial \rho}{\partial T}\right)_{\mu}
		+ \frac{1}{\rho} \left(\frac{\partial \rho}{\partial \mu}\right)_{T}
		\left(\frac{\partial P}{\partial T}\right)_{\mu}
		\left(\frac{\partial P}{\partial \mu}\right)^{-1}_{T}
	\end{equation}
	which is readily calculated based on Eqs.~\eqref{eq:eosMT} and~\eqref{eq:density}.\\
	(NB. Compare this equation with Eqs.(57)-(58) from~\cite{StrokerMeier2021})

	We introduce the reduced thermal expansion coefficient by
	\begin{equation}
		\alpha^*_P = \frac{\varepsilon}{k_{\rm B}} \alpha_P.
	\end{equation}
	Figure~\ref{fig4a} shows the dependency of the reduced thermal expansion coefficient $\alpha^*_P$ on the density $\bar{n}$ for various values of temperature above the critical one. The dependency of $\alpha^*_P$ on the effective chemical potential $M$ is displayed in Figure~\ref{fig4b}.

>>>>

	\subsection{Thermal expansion coefficient}
	The thermal expansion coefficient is defined by
	\begin{equation}
		\alpha_P = \frac{1}{V}\left(\frac{\partial V}{\partial T}\right)_{P,\langle N\rangle}.
	\end{equation}
	We rewrite $\alpha_P$ into a form suitable for the equation of state $P=P(T,\mu)$, see~\eqref{eq:eosMT} and~\eqref{eq:eosPTM_reduced}
	\begin{eqnarray}
		\alpha_P & = & \frac{1}{V}\left(\frac{\partial V}{\partial T}\right)_{P, N}
		\nonumber\\
		\label{eq:alpha_vs_rho}
		& = & -\frac{1}{\rho} \left(\frac{\partial \rho}{\partial T}\right)_{P, N}
		\\
		& = & -\frac{1}{\rho} \left(\frac{\partial \rho}{\partial T}\right)_{\mu} 
		 - \frac{1}{\rho} \left(\frac{\partial \rho}{\partial \mu}\right)_{T}
		 \left(\frac{\partial \mu}{\partial T}\right)_{P, N}.
	\end{eqnarray}
	Applying the cyclic relation
	\begin{equation*}
		\left(\frac{\partial \mu}{\partial T}\right)_{P, N}
		\left(\frac{\partial T}{\partial P}\right)_{\mu, N}
		\left(\frac{\partial P}{\partial \mu}\right)_{T, N}
		= -1
	\end{equation*}
	we get
	\begin{eqnarray*}
		\left(\frac{\partial \mu}{\partial T}\right)_{P, N} & = & 
		- \left(\frac{\partial P}{\partial T}\right)_{\mu}
		\left(\frac{\partial P}{\partial \mu}\right)^{-1}_{T}.
	\end{eqnarray*}
	Substituting the result into the expression for $\alpha_P$, we arrive at the final formula for the thermal expansion coefficient
	\begin{equation}
		\alpha_P = -\frac{1}{\rho}\left(\frac{\partial \rho}{\partial T}\right)_{\mu}
		+ \frac{1}{\rho} \left(\frac{\partial \rho}{\partial \mu}\right)_{T}
		\left(\frac{\partial P}{\partial T}\right)_{\mu}
		\left(\frac{\partial P}{\partial \mu}\right)^{-1}_{T}
	\end{equation}
	which is readily calculated based on Eqs.~\eqref{eq:eosMT} and~\eqref{eq:density}.\\
	(NB. Compare this equation with Eqs.(57)-(58) from~\cite{StrokerMeier2021})

	We introduce the reduced thermal expansion coefficient by
	\begin{equation}
		\alpha^*_P = \frac{\varepsilon}{k_{\rm B}} \alpha_P.
	\end{equation}
	It is expressed in terms of the reduced quantities $P^*$ and $\bar{n}$ as follows
	\begin{equation}
		\alpha^*_P = \varepsilon \beta_c 
		\left( 
		-\frac{1}{\bar{n}}\left(\frac{\partial \bar{n}}{\partial \tau}\right)_{\mu}
		+ \frac{1}{\bar{n}} \left(\frac{\partial \bar{n}}{\partial \mu}\right)_{T}
		\left(\frac{\partial P^*}{\partial \tau}\right)_{\mu}
		\left(\frac{\partial P^*}{\partial \mu}\right)^{-1}_{T} 
		\right).
	\end{equation}

	If it is preferable to use the equation of state in the form $P=P(T, \bar{n})$, see~\eqref{eq:eosNT} and~\eqref{eq:eosPTn_reduced}, then we need to rewrite the definition for $\alpha_P$ accordingly. We start with~\eqref{eq:alpha_vs_rho} and apply the cyclic relation
	\begin{equation}
		\left(\frac{\partial \rho}{\partial T}\right)_P = - \left(\frac{\partial P}{\partial T}\right)_{\rho} \left(\frac{\partial P}{\partial \rho}\right)^{-1}_T
	\end{equation}
	to get for $\alpha_P$
	\begin{equation}
		\alpha_P = \frac{1}{\rho} \left(\frac{\partial P}{\partial T}\right)_{\rho} \left(\frac{\partial P}{\partial \rho}\right)^{-1}_T.
	\end{equation}
	or for $\alpha^*_P$
	\begin{equation}
		\alpha^*_P = \frac{\varepsilon\beta_c}{\bar{n}} \left(\frac{\partial P^*}{\partial \tau}\right)_{\rho} \left(\frac{\partial P^*}{\partial \bar{n}}\right)^{-1}_T.
	\end{equation}

	Figure~\ref{fig4a} shows the dependency of the reduced thermal expansion coefficient $\alpha^*_P$ on the density $\bar{n}$ for various values of temperature above the critical one. The dependency of $\alpha^*_P$ on the effective chemical potential $M$ is displayed in Figure~\ref{fig4b}.
